Neural Circuit Organization of the Ctenophore Apical Organ

### 2025-03-20 manuscript/Manuscript.qmd
MS: edits

### Identification of Synaptic Structures in Syncytium Neurons

@@

Furthermore, ANNs are also different from other neurons previously reported by EM, including mesogleal neurons and ciliated sensory cells that synalse on the SNN (types 1-4)[@Burkhardt2023]. Based on their distinct size and position, we classified the ANNs into two types. The first type is a single large neuron (ANN_Q1-4) that has four (possibly five) nuclei and spans all four quadrants (Q1-Q4). This neuron has X <!--# add number --> presynaptic sites. The second type consists of two neurons (ANN_Q1Q2 and ANN_Q3Q4), each with two nuclei spanning two quadrants, and each containing X presynaptic structures.
<!--#  -->

![**Figure 2. Identification of Three Syncytial Neurons in the Apical Organ**\
(A) 3D reconstruction of the aboral nerve net (ANN) Q1Q2Q3Q4 (Q1-4), which spans all four quadrants and contains multiple nuclei (blue). The spheres represent the positions of individual nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(B) 3D reconstruction of two ANN Q1Q2 (pink) and Q3Q4 (orange), each spanning two quadrants and containing multiple nuclei. The spheres indicate the positions of the nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(C) Graph showing the number of mitochondria per cell and the proportion of mitochondria involved in synapse formation for each cell type. Green represents mitochondria associated with synapses, while gray represents mitochondria not involved in synapse formation. ANN has a significantly higher number of mitochondria compared to other cell types, with most of them forming synapses. The abbreviations are as follows: bal (balancer), brg (bridge), bsl (bristle), cg (ciliated groove), dv (dense vesicle cells), imc (intra-multiciliated cells), la (lamellate bodies), li (lithocytes), pl (plumose), and ef (epithelial floor cells).\
(D) Localization of mitochondria within ANN Q1-4. Red indicates mitochondria associated with the presynaptic triad structures, yellow marks mitochondria containing synaptic vesicles but lacking a clearly defined presynaptic triad, blue represents mitochondria with unclear synaptic vesicles, and black denotes mitochondria where no synaptic vesicles were identified.\
(E) Representative electron micrographs of presynaptic triad structures observed in the dataset and their schematic diagrams. The left diagram illustrates synaptic projections from the center ANN (ANN Q1-4) to the lateral ANN (ANN Q3Q4), while the right diagram shows synaptic projections from the lateral ANN (ANN Q1Q2) to the center ANN (ANN Q1-4). The abbreviations are as follows: mi (mitochondrion), er (endoplasmic reticulum), and sv (synaptic vesicles).\
(F) Plot showing the three-dimensional positions of synapses. Synapses from the center ANN to the lateral ANN are shown in magenta, while synapses from the lateral ANN to the center ANN are shown in cyan. Blue dots indicate the locations of autapses within ANN Q1-4.\
(G) Neural circuit map of ANN connectivity. The blue, orange, and pink circles represent individual neurons in ANN Q1-4, Q2Q3, and Q3Q4, respectively. Magenta arrows indicate synaptic connections from the center ANN to the lateral ANN, while cyan arrows indicate connections from the lateral ANN to the center ANN. Numbers denote the number of synapses. Blue arrows represent autapses in ANN Q1-4 along with their synapse counts.](figures/Figure2.png)

### Identification the Gravity-Sensitive Neural Circuit via the Syncytial Neurons Network

We discovered that ANN neurons form synaptic connections with each other, and that ANN_Q1-4 forms self-synapses, or “autapses.” To our knowledge, previous reports have not identified synaptic connections between subepithelial nerve net (SNN) neurons. Thus, our results represent the first report of synaptic connectivity between neurons in the apical organ, forming a network in ctenophores (subject to confirmation). These ANN networks were found to form many presynaptic structures in relation to the gravity-sensing balancer cells.

Balancer cells are monociliated cells, and their cilia protrude longer than those of other monociliated cells. Moreover, the cilia of balancer cells are bundled into four groups at the center of the apical organ, forming a compound cilium. Based on these features, we identified and classified the balancer cells. The cellular arrangement differed clearly when viewed in the lateral view, sagittal plane, and tentacular plane, with the cell bodies gathering toward the apical organ in the tentacular plane. In each quadrant, the number of cells was as follows: Q1: 37, Q2: 32, Q3: 32, Q4: 28. Each cell contained 3 to 10 ? mitochondria.

>>>>

### An aboral synaptic nerve net of syncytial neurons

@@

Furthermore, ANNs are also different from other neurons previously reported by EM, including mesogleal neurons and ciliated sensory cells that synalse on the SNN (types 1-4)[@Burkhardt2023]. Based on their distinct size and position, we classified the ANNs into two types. The first type is a single large neuron (ANN_Q1-4) spans all four quadrants (Q1-Q4) and has four (possibly five) nuclei This neuron has X <!--# add number --> presynaptic sites. The second type consists of two neurons (ANN_Q1Q2 and ANN_Q3Q4), each spanning two quadrants and with two nuclei, and each containing X presynaptic structures.
<!--#  -->

![**Figure 2. Organisation of the aboral synaptic nerve net**\
(A) 3D reconstruction of the aboral nerve net (ANN) Q1Q2Q3Q4 (Q1-4), which spans all four quadrants and contains multiple nuclei (blue). The spheres represent the positions of individual nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(B) 3D reconstruction of two ANN Q1Q2 (pink) and Q3Q4 (orange), each spanning two quadrants and containing multiple nuclei. The spheres indicate the positions of the nuclei. The left panel shows a aboral view of the apical organ, the middle panel presents a sagittal plane view, and the right panel provides a tentacular plane view.\
(C) Graph showing the number of mitochondria per cell and the proportion of mitochondria involved in synapse formation for each cell type. Green represents mitochondria associated with synapses, while gray represents mitochondria not involved in synapse formation. ANN has a significantly higher number of mitochondria compared to other cell types, with most of them forming synapses. The abbreviations are as follows: bal (balancer), brg (bridge), bsl (bristle), cg (ciliated groove), dv (dense vesicle cells), imc (intra-multiciliated cells), la (lamellate bodies), li (lithocytes), pl (plumose), and ef (epithelial floor cells).\
(D) Localization of mitochondria within ANN Q1-4. Red indicates mitochondria associated with the presynaptic triad structures, yellow marks mitochondria containing synaptic vesicles but lacking a clearly defined presynaptic triad, blue represents mitochondria with unclear synaptic vesicles, and black denotes mitochondria where no synaptic vesicles were identified.\
(E) Representative electron micrographs of presynaptic triad structures observed in the dataset and their schematic diagrams. The left diagram illustrates synaptic projections from the center ANN (ANN Q1-4) to the lateral ANN (ANN Q3Q4), while the right diagram shows synaptic projections from the lateral ANN (ANN Q1Q2) to the center ANN (ANN Q1-4). The abbreviations are as follows: mi (mitochondrion), er (endoplasmic reticulum), and sv (synaptic vesicles).\
(F) Plot showing the three-dimensional positions of synapses. Synapses from the center ANN to the lateral ANN are shown in magenta, while synapses from the lateral ANN to the center ANN are shown in cyan. Blue dots indicate the locations of autapses within ANN Q1-4.\
(G) Neural circuit map of ANN connectivity. The blue, orange, and pink circles represent individual neurons in ANN Q1-4, Q2Q3, and Q3Q4, respectively. Magenta arrows indicate synaptic connections from the center ANN to the lateral ANN, while cyan arrows indicate connections from the lateral ANN to the center ANN. Numbers denote the number of synapses. Blue arrows represent autapses in ANN Q1-4 along with their synapse counts.](figures/Figure2.png)

### Synaptic connectome of the gravisensory organ

Our synapse annotation revealed synapses between the ANN neurons and synapses that ANN_Q1-4 forms on itsels (autapses). The ANN neurons also form synapses on the gravity-sensing balancer cells.

Balancer cells are monociliated cells with long motile cilia that form four bundles of compound cilia, one in each quadrant. These bundles come together at the centre of the apical organ to support the cellular mass of the statolyth. The bending of the cilia and the position of somata differed clearly when viewed laterally from the sagittal or the tentacular plane (Figure 3B). In each quadrant, there were between 28-37 balancer cells (Q1: 37; Q2: 32; Q3: 32; Q4: 28). Each cell contained 3 to 10 <!-- check --> mitochondria.
